---
title: "AnVILBestPractices"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{AnVILBestPractices}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Illustration of analytic workspace

We'll use [celldex](https://bioconductor.org/packages/celldex)
to examine expression in bulk RNA-seq data from
human primary cells.  These are used to form signatures
for the classification of single cell types.

First we get an overview of variation in expression
patterns in a reference dataset using principal components.
```{r lkcd, message=FALSE}
library(AnVILBestPractices)
library(SummarizedExperiment)
library(irlba)
hp = celldex::HumanPrimaryCellAtlasData()
types = c("B_cell", "iPS_cells", "Tissue_stem_cells", 
  "Monocyte", "Endothelial_cells", "T_cells", "DC", 
  "Macrophage")
hplim = hp[, which(hp$label.main %in% types)]
pc = prcomp_irlba(t(assay(hplim)), n=10)
pairs(pc$x[,1:5], col=factor(hplim$label.main), cex=.3,
   pch=19)

```

Now we use the reference to classify cells in
a PBMC dataset, using the SingleR methodology.

```{r doclass, message=FALSE}
p3k = TENxPBMCData::TENxPBMCData("pbmc3k")
rownames(p3k) = make.names(rowData(p3k)$Symbol, unique=TRUE)
p3k = scater::logNormCounts(p3k)
stats <- scran::modelGeneVar(p3k)
hvg = scran::getTopHVGs(stats, n=5000)
pred3k = do_SingleR(p3k[hvg,], BPPARAM=BiocParallel::MulticoreParam())
head(pred3k$preds)
```

Check the separability of asserted classes in PCA space:
```{r dopca}
pc = prcomp_irlba(t(assay(pred3k$input)), 10)
pairs(pc$x[,1:5], pch=19, cex=.3, col=factor(pred3k$preds$pruned.labels))
```
